---
title: "Advanced ML Optimization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced ML Optimization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

The `MLSampling` package introduces advanced machine learning capabilities for spatial sampling design.
This vignette explores two powerful new methods: Bayesian Deep Learning (BDL) for uncertainty-aware sampling
and Random Forest (RF) optimization for feature-driven design. It also covers how to combine these approaches
using Ensemble methods.

# Bayesian Deep Learning (BDL)

BDL allows us to quantify the uncertainty in our spatial predictions. By sampling in areas of high uncertainty,
we can maximize the information gain of new samples.

## Epistemic vs Aleatoric Uncertainty

- **Epistemic Uncertainty**: Uncertainty due to lack of knowledge (model uncertainty). Can be reduced with more data.
- **Aleatoric Uncertainty**: Uncertainty due to inherent noise in the data. Cannot be reduced with more data.

The `MLSampling` tool typically targets **Epistemic** or **Total** uncertainty for sampling optimization.

## Running BDL

```{r}
# Run BDL optimization
bdl_result <- tool$run_bdl(
  field_data = field_data,
  existing_samples = existing_samples,
  n_new_samples = 30,
  uncertainty_type = "epistemic",  # Target model uncertainty
  mc_iterations = 100              # Number of Monte Carlo passes
)

# Visualize uncertainty map
tool$visualization_service$plot_uncertainty_map(
  uncertainty_raster = bdl_result$uncertainties$epistemic_uncertainty,
  field_data = field_data,
  title = "Epistemic Uncertainty Map"
)
```

# Random Forest Optimization (RF)

Random Forest is excellent for understanding non-linear relationships between covariates and the target variable.
The RF optimization module uses feature importance to weigh the sampling design.

## Spatial Random Forest

We incorporate spatial autocorrelation by adding spatial lag features to the model.

```{r}
rf_result <- tool$run_rf_optimization(
  field_data = field_data,
  existing_samples = existing_samples,
  n_new_samples = 30,
  feature_importance_method = "permutation",
  spatial_autocorr = TRUE  # Enable spatial features
)

# View feature importance
print(rf_result$feature_importance)
```

# Ensemble Optimization

Why choose one model when you can use them all? The Ensemble Manager combines predictions and sampling suggestions
from multiple models (BDL, RF, UDL, UFN) to create a robust design.

## Voting vs Stacking

- **Voting**: Each model "votes" for locations, and the most popular locations are selected.
- **Stacking**: A meta-model learns to combine the predictions of base models.

```{r}
ensemble_result <- tool$run_ensemble(
  field_data = field_data,
  existing_samples = existing_samples,
  n_new_samples = 50,
  methods = c("BDL", "RF", "UDL"),
  ensemble_method = "voting"
)

# Generate ensemble report
tool$generate_ml_report(ensemble_result)
```

# Conclusion

By leveraging these advanced ML techniques, `MLSampling` moves beyond simple geometric or statistical sampling designs, allowing for:
1. **Risk-aware sampling** (via Uncertainty Quantification)
2. **Process-aware sampling** (via Feature Importance)
3. **Robust sampling** (via Ensemble methods)
